\documentclass[10pt]{article}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{amsfonts}
\usepackage{listings}
\author{BM Corser}
\title{Algebra 2 Assignment 1}
\newcommand*{\E}{$e$}
\newcommand*{\A}{$\alpha$ }
\newcommand*{\As}{$\alpha^2$ }
\newcommand*{\Ac}{$\alpha^3$ }
\newcommand*{\B}{$\beta$ }
\newcommand*{\AB}{$\alpha\beta$ }
\newcommand*{\AsB}{$\alpha^2\beta$ }
\newcommand*{\AcB}{$\alpha^3\beta$ }
\newcommand*{\Zp}{\mathbb{Z}_p }
\begin{document}
  \maketitle 
  \begin{enumerate}
    \item
    \begin{enumerate}
        \item Let \A represent a clockwise rotation of $\frac{\pi}{2}$ about
          the origin and $\beta$ represent a particular reflection. Now the
          Cayley table of ${\rm Dih}(8)$ can be written

        \begin{tabular}{r|cccccccc}
               & \E   & \A   & \As  & \Ac  & \B   & \AB  & \AsB & \AcB \\ \hline
          \E   & \E   & \A   & \As  & \Ac  & \B   & \AB  & \AsB & \AcB \\
          \A   & \A   & \As  & \Ac  & \E   & \AB  & \AsB & \AcB & \B   \\
          \As  & \As  & \Ac  & \E   & \A   & \AsB & \AcB & \B   & \AB  \\
          \Ac  & \Ac  & \E   & \A   & \As  & \AcB & \B   & \AB  & \AsB \\
          \B   & \B   & \AcB & \AsB & \AB  & \E   & \Ac  & \As  & \A   \\
          \AB  & \AB  & \B   & \AcB & \AsB & \A   & \E   & \Ac  & \As  \\
          \AsB & \AsB & \AB  & \B   & \AcB & \As  & \A   & \E   & \Ac  \\
          \AcB & \AcB & \AsB & \AB  & \B   & \Ac  & \As  & \A   & \E   \\
        \end{tabular}
      \item The set of elements that commute with every other element in Dih(8)
        is $\{\alpha^2, e\}$ this can be observed in the Cayley table above by
        noticing that for each of these elements, both the row and column for
        that element contain the same elements in the same order.
      \item The subgroups of Dih(8) are the trivial and identity subgroups
        $\{e\}$ and Dih(8), the four 2-subgroups formed by every reflecting
        element and the identity, $\{e, \alpha^n\beta\}$ for $0 \leq n \leq 3$,
        the commutative 2-subgroup $\{e, a^2\}$, the 4-subgroup containing the
        identity and all rotation elements $\{e, \alpha, \alpha^2, \alpha^3\}$
        and finally the 4-subgroup containing $\{e, \alpha^2, \alpha\beta,
        \alpha^3\beta\}$.

        By Lagrange, the order of any subgroup $H \leq {\rm Dih}(8)$, $|H|$
        must divide $|{\rm Dih}(8)|$. So there aren't any 3- or 5-subgroups
        because neither 3 or 5 divide 8.

        Now, a subgroup must use a subset of Dih(8) and we needn't consider
        subsets of size 3 or 5.

        There can only be one 1-subgroup, $\{e\}$, so there aren't any more of
        those.

        There aren't any 2-subgroups beyond those already mentioned, because
        none of the non-reflecting 2-subsets $\{e, \alpha^n\}$ with $1 \leq n
        \leq 3$ have closure.

        There aren't any more 4-subgroups because (with the expection of the
        subgroup $\{e, \alpha^2, \alpha\beta, \alpha^3\beta\}$) any subgroup
        that has a combination of reflections and rotations doesn't have
        closure.

        There can only be one 8-subgroup, Dih(8), so there aren't any more of
        those either.
    \end{enumerate}
  \item Let $k, l \in K$ and $1_K$ be the identity element of $K$.
    Since $K \leq G$, we know
    \begin{align*}
      k, l, kl \in G.
    \end{align*}
    Because $f : G \rightarrow H$, certainly 
      $f(k), f(l), f(kl) \in H$ and $f(K) \subseteq H$.

    Since $f$ is a homomorphism, we know that $f(ab) = f(a)f(b)$ and therefore
    \begin{align}
      f(kl) = f(k)f(l) \in H,
    \end{align}
    and, if $1_H$ is the identity element of $H$,
    \begin{align}
      f(1_K) = 1_H.
    \end{align}
          The fact that $f$ is a homomorphism also tells us that $f(a^{-1}) =
          f(a)^{-1}$ and since $k^{-1} \in K$ it is also true that
    \begin{align}
      f(k)^{-1} \in H.
    \end{align}
    Due to the facts that $K \leq G$ and that $f$ is a homomorphism, it is also
    true that if $j \in K$,
      \begin{align}
        f((jk)l) = f(jk)f(l) = f(j)f(k)f(l) = f(j)f(kl) = f(j(kl)).
      \end{align}

      Therefore $f(K)$ is closed (1), has an identity element $1_H$ (2), has
      inverses (3) and is associative (4). Hence, $f(K) \leq H$.
    \item 
      \begin{enumerate}
        \item $V_4 = \{e, a, b, c\}$ such that

          $ab = ba = c$,

          $ac = ca = b$,

          $bc = cb = a$ and

          $a^2 = b^2 = c^2 = e$.

          Let $f$ be an automorphism on $V_4$. Because $f$ is a homomorphism,
          $f(e) = e$, that is, $f$ fixes $e$.

          Because of this and because of the definition of $V_4$, any
          permutation on $\{e, a,b,c\}$ where $e$ is fixed is an automorphism
          of which there are $\Big|\{a,b,c\}\Big|! = 6$.
        \item There is clearly one automorphism on $\mathbb{Z}_4$, that is the
          identity automorphism $\theta : \mathbb{Z}_4 \rightarrow
          \mathbb{Z}_4$ where $\theta(n) = n$ for all $n \in \mathbb{Z}_4$.
          There is also the automorphism $\vartheta : \mathbb{Z}_4
          \rightarrow\mathbb{Z}_4$, which for $n \in \mathbb{Z}_4$
          $$
          \vartheta(n) =
            \begin{cases}
              &\vartheta(0) = 0 \\
              &\vartheta(1) = 3 \\
              &\vartheta(2) = 2 \\
              &\vartheta(3) = 1 \\
            \end{cases}
          $$
        \item From the starting point that an automorphism $f$ on $\mathbb{Z}_p$ must be a permutation statisfying the homomorphism
          property $f(ab) = f(a)f(b)$ for all $a, b \in \mathbb{Z}_p$, I wrote
          a short C++ program to test permutations for this property. The code
          is available at the following URL

          \url{https://git.io/vFSef}.

          and its output for $p \in \{5,7,11\}$ was

          \begin{lstlisting}

Z_p, p = 5 has 24 permutations
  ( 0 1 2 3 4 )
  ( 0 2 4 1 3 )
  ( 0 3 1 4 2 )
  ( 0 4 3 2 1 )
Z_p, p = 5 has 4 automorphisms

Z_p, p = 7 has 720 permutations
  ( 0 1 2 3 4 5 6 )
  ( 0 2 4 6 1 3 5 )
  ( 0 3 6 2 5 1 4 )
  ( 0 4 1 5 2 6 3 )
  ( 0 5 3 1 6 4 2 )
  ( 0 6 5 4 3 2 1 )
Z_p, p = 7 has 6 automorphisms

Z_p, p = 11 has 3.6288e+06 permutations
  ( 0 1 2 3 4 5 6 7 8 9 10 )
  ( 0 2 4 6 8 10 1 3 5 7 9 )
  ( 0 3 6 9 1 4 7 10 2 5 8 )
  ( 0 4 8 1 5 9 2 6 10 3 7 )
  ( 0 5 10 4 9 3 8 2 7 1 6 )
  ( 0 6 1 7 2 8 3 9 4 10 5 )
  ( 0 7 3 10 6 2 9 5 1 8 4 )
  ( 0 8 5 2 10 7 4 1 9 6 3 )
  ( 0 9 7 5 3 1 10 8 6 4 2 )
  ( 0 10 9 8 7 6 5 4 3 2 1 )
Z_p, p = 11 has 10 automorphisms

          \end{lstlisting}

          The program suggested a clear result, namely that there are $p - 1$
          many automorphisms $f$ on $\Zp$. By inspecting the output, it is easy
          to observe (for those primes tested) that for all $\{g : \langle g \rangle = \Zp\}$
          and $a \in \Zp$ $$
          f(a) = \begin{cases}
            f(0) = 0 \\
            f(a) = g^a \\
          \end{cases}
          $$

          Because $f$ is a permutation, it is a bijection. Also the
          homomorphism condition $f(ab) = f(a)f(b)$ for all $a, b \in
          \mathbb{Z}_p$ holds
          \begin{align*}
            f(ab)  &= f(a)f(b) \\
            f(a \oplus b)  &= f(a) \oplus f(b) \\
            g^{a \oplus b} &= g^a \oplus g^b \\
                   (a \oplus b) \times g &= (a \times g) \oplus (b \times g)  \text{(due to group operation)} \\
                   &= (a \oplus b) \times g 
          \end{align*}

      \end{enumerate}

  \end{enumerate}
\end{document}
