\documentclass[10pt]{article}
\usepackage{mathtools}
\usepackage{amsfonts}
\newcommand*{\La}{\mathcal{L}}
\newcommand*{\dxdt}[1]{\frac{\text{d}#1}{\text{d}t}}
\newcommand*{\nip}{\left(\tfrac{400}{\pi n}\right)}
\newcommand*{\dx}{\text{d}x}
\newcommand*{\px}[2]{\tfrac{\partial\dot #1}{\partial #2}}

\author{BM Corser}
\title{Calculus 3 Assignment 2}
\date{January 18, 2018}
\begin{document}
  \maketitle
  \begin{enumerate}
    \item Let $\dxdt{x} = y$, now
      \begin{align*}
        (t^2+1)\dxdt{y} - xty + x^2 &= x\cos t \\
        (t^2+1)\dxdt{y} &= x(\cos t + ty - x) \\
        \dxdt{y} &= x\left(\frac{\cos t + ty - x}{t^2+1}\right) \\
      \end{align*}
      let $z = t$. Now we can write the original differential equation as a
      system of first order autonomous differential equations
      \begin{align*}
        \dxdt{x} &= y \\
        \dxdt{y} &= x\left(\frac{\cos z + zy - x}{z^2+1}\right) \\
        \dxdt{z} &= 1.
      \end{align*}
      \pagebreak
    \item $\dot x = \alpha x^2 - x^4 = x^2(\alpha - x^2)$. The fixed points of
      this system are $x = 0$ and $x = \pm \sqrt{\alpha}$. To examine the
      behaviour of the sytem around these fixed points, we introduce an
      ``amount of change'' variable $\sqrt{\delta} \in \mathbb{R}$.

      In the case that $\alpha > 0$, when $x < -\sqrt{\alpha}$ we let
      $\sqrt{\delta} > 1$ and suppose that
      $x = \sqrt{\delta}(-\sqrt{\alpha}) = -\sqrt{\delta\alpha}$, now
      $x^2 = \delta\alpha > \alpha$, and the derivative of $x$,
      $\dot x = \delta\alpha(\alpha - \delta\alpha) < 0$.
      In the case $-\sqrt{\alpha} < x < 0$,
      let $0 < \sqrt{\delta} < 1$ and suppose $x = -\sqrt{\delta\alpha}$,
      now $x^2 = \delta\alpha < \alpha$ and
      $\dot x = \delta\alpha(\alpha-\delta\alpha) > 0$.
      In the case $0 < x < \sqrt{\alpha}$,
      let $0 < \sqrt{\delta} < 1$ and suppose $x = \sqrt{\delta\alpha}$,
      now $x^2 = \delta\alpha < \alpha$ and
      $\dot x = \delta\alpha(\alpha-\delta\alpha) > 0$.
      In the case $x > \sqrt{\alpha}$,
      let $\sqrt{\delta} > 1$ and suppose $x = \sqrt{\delta\alpha}$,
      now $x^2 = \delta\alpha > \alpha$ and
      $\dot x = \delta\alpha(\alpha-\delta\alpha) < 0$. We now have the
      necessary information to write our phase portrait for the system when
      $\alpha > 0$,

      \bigskip
      \bigskip
      \bigskip
      \bigskip
      \bigskip
      \bigskip
      \bigskip
      \bigskip

      It is clear that the fixed point $x = -\sqrt{\alpha}$ is unstable, $x =
      0$ is semistable and $x = \sqrt{\alpha}$ is stable.

      In the case that $\alpha < 0$, since this dynamical system is
      one-dimensional, we know that $x \not \in \mathbb{C}$ so the only fixed
      point is $x = 0$ and our dynamical system can be written $\dot x = -x^4$,
      but our phase portrait still provides a satisfactory illustration when
      the interval between $-\sqrt{\alpha}$ and $\sqrt{\alpha}$ is exactly
      zero. The fixed point $x = 0$ is semistable.


      \item
        \begin{enumerate}
          \item $A = \begin{pmatrix} 3 & 0 \\ \beta & 3 \end{pmatrix}$, $\det(A
              - \lambda I) = (3 - \lambda)^2 = 0$, $\lambda = 3$ with algebraic
              multiplicity 2.

            \item
              \begin{enumerate}
                  \item $\beta = 0$:
                    \begin{align*}
                      (A - \lambda I)\mathbf{x} &= \mathbf{0} \\
                      \begin{pmatrix} 0 & 0 \\ 0 & 0 \end{pmatrix}
                      \begin{pmatrix} x \\ y \end{pmatrix}
                        &=
                      \begin{pmatrix} 0 \\ 0 \end{pmatrix} \\
                    \end{align*}
                  Any value of $x$ and $y$ will satisfy this equation, so the eigenvectors are 
                      $\begin{pmatrix} 1 \\ 0 \end{pmatrix}$ and
                      $\begin{pmatrix} 0 \\ 1 \end{pmatrix}$. Because $A$ is a
                        $2\times 2$ matrix and there are 2 independent
                        eigenvectors, $A$ is not defective when $\beta = 0$.
                  \item $\beta \not= 0$:
                    \begin{align*}
                      (A - \lambda I)\mathbf{x} &= \mathbf{0} \\
                      \begin{pmatrix} 0 & 0 \\ \beta & 0 \end{pmatrix}
                      \begin{pmatrix} x \\ y \end{pmatrix}
                        &=
                      \begin{pmatrix} 0 \\ 0 \end{pmatrix} \\
                      \begin{pmatrix} 0 \\ \beta x \end{pmatrix}
                        &=
                      \begin{pmatrix} 0 \\ 0 \end{pmatrix} \\
                    \end{align*}
                  To satisfy this equation $x$ must be equal to zero, so the only eigenvector of this $A$ is
                      $\begin{pmatrix} 0 \\ 1 \end{pmatrix}$.  Becaues there is
                        only one independent eigenvector, $A$ does not have a
                        complete base of eigenvectors and is defective when
                        $\beta \not= 0$.
              \end{enumerate}
            \item When $\beta = 0$ the fixed point $(0,0)$ can be classified as
              an unstable star, and the phase portrait can be written

              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip

              where eigenvectors are the axes of the graph.

              When $\beta = 3$ the fixed point $(0, 0)$ can be classified as an
              unstable improper node and the phase portrait can be written

              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip
              \bigskip

              where the single eigenvector is the $y$-axis of the graph.


      \item Because the system has equal eigenvalues, $\lambda_1 = \lambda_2 = \lambda$, the general form
        \begin{align*}
          \mathbf{x}(t) &= c_1e^{\lambda_1t}\mathbf{x}_1+c_2e^{\lambda_2t}\mathbf{x}_2 \\
                        &= e^{\lambda t}(c_1\mathbf{x}_1+c_2\mathbf{x}_2).
        \end{align*}
        Substituting values for the initial conditions of $\mathbf{x}(t)$ when
        $t = 0$, $\lambda = 3$ and eigenvectors $\mathbf{x}_1$ and
        $\mathbf{x}_2$ we can write
        \begin{align*}
          \begin{pmatrix} 2 \\ -1 \end{pmatrix} &=
            e^{3 \cdot 0}\left(c_1\begin{pmatrix} 1 \\ 0 \end{pmatrix} +
              c_2\begin{pmatrix} 0 \\ 1 \end{pmatrix}\right) \\
                &= \begin{pmatrix} c_1 \\ 0 \end{pmatrix} +
              \begin{pmatrix} 0 \\ c_2 \end{pmatrix} \\
                &= \begin{pmatrix} c_1 \\ c_2 \end{pmatrix} \\
        \end{align*}
        therefore $c_1 = 2$, $c_2 = -1$ and
        $\mathbf{x}(t) = e^{3t}\begin{pmatrix} 2 \\ -1 \end{pmatrix}$.
        \end{enumerate}
      \item The dynamical system in the question can be written
        \begin{align*}
          \dot x &= y \\
          \dot y &= -4\sin x - 3y.
        \end{align*}
        The fixed points of this system are those that satisfy
        $\dot x = \dot y = 0$. We can write $y = 0$ and see that
        $-4\sin x = 0$. The latter is true when $x$ takes values in
        $\{n\pi : n \in \mathbb{Z}\}$. Therefore the fixed points of this
        system are $\{(n\pi,0) : n \in \mathbb{Z}\}$.
        Computing the Jacobian of the system gives
        \begin{align*}
          J = \begin{pmatrix}
            \px{x}{x} & \px{x}{y} \\
            \px{y}{x} & \px{y}{y} \\
          \end{pmatrix} &=
          \begin{pmatrix}
            0 & 1 \\
            -4\cos x & -3 \\
          \end{pmatrix}.
        \end{align*}
      Given that values of $x$ are $n\pi$, we can see that there are two
      possibilities for the bottom left entry $-4\cos x$, namely, $4$ when
      $n$ is odd and $-4$ when $n$ is even.

      Evaluating $J$ at the fixed point $(n\pi,0)$ when $n$ is odd, gives the
      Jacobian matrix
        $
          J_\pi = \begin{pmatrix}
            0 & 1 \\
            4 & -3 \\
          \end{pmatrix}
        $.

        That 
        $$\det(J_\pi - \lambda I) = -\lambda(-3-\lambda) - 4 = (\lambda + 4)(\lambda - 1)$$
        tells us that the eigenvalues for $J_\pi$ are $\lambda
        = -4$ and $\lambda = 1$. Since these eigenvalues both have a nonzero
        real part, we know that the fixed points $\{(n\pi,0):n\text{ is odd}\}$ are
        hyperbolic and we can apply Hartman-Grobman.

        For $\lambda = -4$,
        \begin{align*}
          (J_\pi - \lambda I)\mathbf{x} &= \mathbf{0} \\
          \begin{pmatrix}
            4 & 1 \\
            4 & 1 \\
          \end{pmatrix}
          \begin{pmatrix}
            x \\
            y \\
          \end{pmatrix}
          &=
          \begin{pmatrix}
            0 \\
            0 \\
          \end{pmatrix}
        \end{align*}
        and the eigenvectors associated with $\lambda = 4$ are of the form $
          \begin{pmatrix}
            4 \\
            -1 \\
          \end{pmatrix}$.

        For $\lambda = 1$,
        \begin{align*}
          (J_\pi - \lambda I)\mathbf{x} &= \mathbf{0} \\
          \begin{pmatrix}
            -1 & 1 \\
            4 & -1 \\
          \end{pmatrix}
          \begin{pmatrix}
            x \\
            y \\
          \end{pmatrix}
          &=
          \begin{pmatrix}
            0 \\
            0 \\
          \end{pmatrix}
        \end{align*}
        and the eigenvectors associated with $\lambda = 1$ are of the form $
          \begin{pmatrix}
            1 \\
            1 \\
          \end{pmatrix}$. The general form of the linearised system in the
          neighbourhood of fixed points $\{(n\pi,0):n \text{ is odd}\}$ can
          therefore be written
          \begin{align*}
            \mathbf{x}(t) &= Ae^{\lambda_1t}\mathbf{x}_1+Be^{\lambda_2t}\mathbf{x}_2 \\
             &= Ae^{-4t}
          \begin{pmatrix}
            4 \\
            -1 \\
          \end{pmatrix}
          + Be^t
          \begin{pmatrix}
            1 \\
            1 \\
          \end{pmatrix}.
          \end{align*}
          Because eigenvalues are real and with opposite signs we can say that,
          locally, these fixed points are saddle points with
          $\mathbf{x}_1$ pushing trajectories inwards and
          $\mathbf{x}_2$ pushing trajectories outwards. Sending $t \rightarrow
          -\infty$ tells us that the $\mathbf{x}_1$ dominates early
          times and $t \rightarrow \infty$ tells us $\mathbf{x}_2$ dominates
          late times.

      Evaluating $J$ at the fixed point $(n\pi,0)$ when $n$ is even, gives the
      Jacobian matrix
        $
          J_0 = \begin{pmatrix}
            0 & 1 \\
            -4 & -3 \\
          \end{pmatrix}
        $.

        $$\det(J_0 - \lambda I) = -\lambda(-3-\lambda) + 4$$ doesn't
        factorise, which tells us $\lambda \in \mathbb{C}$. Applying the
        quadratic equation,
        $\lambda = \frac{-3\pm\sqrt{(-3)^2-4\cdot 1 \cdot 4}}{2 \cdot 1} = \tfrac{1}{2}(-3 \pm i\sqrt{7})$.
        These complex conjugate values for $\lambda$ both have a nonzero real
        part, so this fixed point is also hyperbolic and we can apply
        Hartman-Grobman.

        We don't need to caculate the eigenvectors in this case because we know
        these fixed points behave like a spirals locally. Because
        $\operatorname{Re}(\lambda) < 0$ we also know that these spirals are
        stable. As for the winding direction, because $\dot x = y$ and $\dot y
        = -3y$ when $x = 0$ these spirals wind in a clockwise direction.

        \pagebreak

        One last thing to consider before writing the phase portrait of this
        system is the interaction between the two classes of fixed points with
        respect to a given trajectory. We have establised that for those fixed
        points with $n$-odd, $\mathbf{x}_1$ dominates early times and
        $\mathbf{x}_2$ dominates late times. For fixed points with $n$-even,
        the idea of early/late times doesn't really make sense, since the
        conjugate eigenvalues mean the eigenvectors ``share'' dominance over
        time, but always in a spiral inwards. In that sense, fixed points with
        $n$-even have a constant amount of ``spiralling dominance'' over the
        trajectory. To get an idea of what happens when all the fixed points
        have most similar amount of influence on the trajectory, we consider
        what happens at \emph{mid times} with, let's say, $t=0$.

        At mid times, then, the coefficients of the eigenvectors $\mathbf{x}_1$
        and $\mathbf{x}_2$ are $1$. Between these two (very roughly, not
        considering $c_1$ or $c_2$) the vector with the larger magnitude will
        be more influential, that is $\mathbf{x}_1$. However, the spiral fixed
        point has exactly the same amount of ``spiralling dominance'' as it did
        when $t \rightarrow \pm \infty$, so there is more ``sum dominance''
        over a trajectory that would cause it to behave as if it were being
        attracted to a fixed point.

        This is an attempt to explain why trajectories don't follow
        $\mathbf{x}_2$ vectors as $t \rightarrow \infty$. A more informal
        argument would be that this dynamical system describes a pendulum, and
        pendulums come to rest.

        The phase portrait is written overleaf.

  \end{enumerate}
\end{document}
